---
title: "R Notebook"
output: html_notebook
---

```{r load libraries}
library(tidyverse)
```

```{r}
# read in data

seq.index <- read_delim("~/LabProjects/1000GenomesDUF1220-variationRD/data/1000Genomes_HiSeq.txt",
                        delim = "\t")
colnames(seq.index) <- tolower(colnames(seq.index))

pop.groups <- read.delim("~/LabProjects/ReadDepth/data/1000Genomes/PopulationGroups.csv", sep = ",")
colnames(pop.groups) <- tolower(colnames(pop.groups))

seq.index <- merge(seq.index,pop.groups)
```

```{r}
#Filter on number of reads

samples.100M.mapped <- seq.index %>% group_by(sample_name, center_name, population, pop.group) %>%
  summarise(tot.reads = sum(read_count)) %>%
  filter(tot.reads >= 100e6)
```

## Calculate number each population by sequencing center
```{r}
pop.counts.by.center <- samples.100M.mapped %>%
  group_by(center_name, population) %>%
  summarise(count = n()) 
```

### Visualize data
```{r}
ggplot(pop.counts.by.center, aes(x = center_name, y = count, fill = population)) +
  geom_bar(stat = "identity")
```


```{r}
pops.20.or.more.by.center <- pop.counts.by.center %>% filter(count >= 20)

ggplot(pops.20.or.more.by.center, aes(x = center_name, y = count, fill = population)) +
  geom_bar(stat = "identity")
```

Look at distributions by population group, rather than the smaller populations
```{r}
pop.groups.by.center <- samples.100M.mapped %>%
  group_by(center_name, pop.group) %>%
  summarise(count = n()) 
```

Visualize results
```{r}
ggplot(pop.groups.by.center, aes(x = center_name, y = count, fill = pop.group, label = count)) +
  geom_bar(stat = "identity") +
  geom_text(size = 2, position = position_stack(vjust = 0.5))
```

Randomly select 20 samples from each population where 20 or more samples were sequenced at BGI
```{r}
pops.20.or.more.bgi <- pops.20.or.more.by.center %>% filter(center_name == "BGI")

by.pop.bgi <- samples.100M.mapped %>% 
  filter(center_name == "BGI",
         population %in% pops.20.or.more.bgi$population) %>%
  group_by(population)

rand.20.per.pop.bgi <- sample_n(by.pop.bgi,20)

length(rand.20.per.pop.bgi$sample_name)

rand.20.per.pop.group.rep <- rand.20.per.pop.bgi %>% group_by(pop.group) %>% summarise(count = n())
ggplot(rand.20.per.pop.group.rep, aes(x = pop.group, y = count)) + geom_bar(stat = "identity")
```
Produces a list of 160 samples


Randomly select 20 samples from each population where 20 or more samples were sequenced at SC
```{r}
pops.20.or.more.sc <- pops.20.or.more.by.center %>% filter(center_name == "SC")

by.pop.sc <- samples.100M.mapped %>% 
  filter(center_name == "SC",
         population %in% pops.20.or.more.sc$population) %>%
  group_by(population)

rand.20.per.pop.sc <- sample_n(by.pop.sc,20)

rand.20.per.pop.group.rep <- rand.20.per.pop.sc %>% group_by(pop.group) %>% summarise(count = n())

length(rand.20.per.pop.sc$sample_name)

ggplot(rand.20.per.pop.group.rep, aes(x = pop.group, y = count)) + geom_bar(stat = "identity")
```
Produces a list of 220 samples



